Towards Compiler-Agnostic Performance in Finite-Difference Codes

### 2015-08-07 docs/papers/shallow_performance/shallow_perf.tex
Add description of compiler flags

%Proliferation of lightweight cores is resulting in increased
%complexity in the programming models required to achieve performance.
%Combine with the complexity of e.g. a model simulating circulation in
%the global ocean. Significant challenge in software engineering.
%Propose an approach for meeting this challenge by separating the
%computational science demands from the {\it e.g.} oceanographic
%demands.

@@

Clearly the separation of code into distinct layers may have an effect
on performance. This overhead, and how to get back to the performance
of a hand optimised code, and potentially improve on it, will be
discussed in the remainder of this paper.

@@

As with any full oceanographic model, boundary conditions must be
applied at the edges of the model domain. In the case of Shallow we
have periodic boundary conditions in both the $x$ and $y$ dimensions.
These are simply implemented by having additional (`halo')
rows/columns at the edges of the domain and ensuring the data in them
is up-to-date before it is read. These rows/columns are updated by
copying data from the corresponding row/column on the opposite side of
the domain. In the current work, the lines of code to do these copies
is manually inserted in the PSy-layer routine as required.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methodology(?)}

@@

In Table~\ref{TABLE_compiler_flags} we give the optimisation flags
used with each compiler. These flags are particularly important for
the \psykal versions of the code. The performance of the original
version of the code is much less sensitive to the compiler
options. This is because it consists of a single source file
containing effectively a single routine (excluding those for I/O).

@@

This work made use of the ARCHER UK National Supercomputing Service
(\url{http://www.archer.ac.uk}). We acknowledge use of the Hartree
Centre IBM Power 8 system in this work. The STFC Hartree Centre is a
research collaboratory in association with IBM providing High
Performance Computing platforms funded by the UK's investment in
e-Infrastructure. The Centre aims to develop and demonstrate next
generation software, optimised to take advantage of the move towards
exa-scale computing. This work was funded by the NERC GOcean project,
grant number NE/L012111/1.

>>>>

@@

Clearly the separation of code into distinct layers may have an effect
on performance. This overhead, how to get back to the performance
of a hand-optimised code, and potentially improve on it, will be
discussed in the remainder of this paper.

@@

As with any full oceanographic model, boundary conditions must be
applied at the edges of the model domain. In the case of Shallow we
have periodic boundary conditions in both the $x$ and $y$ dimensions.
These are simply implemented by having additional (`halo')
rows/columns at the edges of the domain and ensuring the data in them
is up-to-date before it is read. These rows/columns are updated by
copying data from the corresponding row/column on the opposite side of
the domain. In the current work, the lines of code to do these copies
are manually inserted in the PSy-layer routine as required.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methodology}

@@

In Table~\ref{TABLE_compiler_flags} we give the optimisation flags
used with each compiler. These flags are particularly important for
the \psykal versions of the code. The performance of the original
version of the code is much less sensitive to the compiler
options. This is because it consists of a single source file
containing effectively a single routine (excluding those for I/O).
For all compilers apart from Cray it was important to use flags that
encourage in-lining. With the Cray compiler some benefit was obtained
by increasing the level of Inter-Procedural Analysis and using
'whole-program optimisation'. In order to get performant code from the
IBM compiler it was found essential to make the pre-fetching
considerably more aggressive than the default.

@@

This work made use of the ARCHER UK National Supercomputing Service
(\url{http://www.archer.ac.uk}). We acknowledge use of the Hartree
Centre IBM Power 8 system in this work. The STFC Hartree Centre is a
research collaboratory in association with IBM providing High
Performance Computing platforms funded by the UK's investment in
e-Infrastructure. The Centre aims to develop and demonstrate next
generation software, optimised to take advantage of the move towards
exa-scale computing. This work was funded by the NERC `GOcean' project,
grant number NE/L012111/1.
